---
title: "JHUDSS C2 R Programming"
author: "mdzakyjaya"
date: '2022-08-02'
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Week 1

## Matriks

pembuatan matriks dapat menggunakan pendefinisan dimensi dengan vector

```{r}
m <- 1:10
dim(m) <- c(2,5)
m
```

pembuatan matriks bisa juga menggunakan funsi cbind() dan rbinf()

```{r}
x <- c(1,3,4)
y <- c(3,4,1)
cmatriks <- cbind(x,y) # menyusun berdasarkan kolom
rmatriks <- rbind(x,y) # menyusun berdararkan baris
```

penamaan dimensi matriks

```{r}
m <- matrix(c(1,2,3,4), nrow = 2, ncol = 2)
dimnames(m) = list(c('a','b'),c('c','d'))
m
```

## Factors

```{r}
x <- factor(c('laki', 'perempuan', 'laki', 'laki', 'perempuan'))
x
table(x) # membuat tabel berisi frequensi dari masing2 faktor
unclass(x) # melihat faktor dari masing value
```

R will order string alpabhetically if ordered levels not predefined.\
defining factors with ordered levels

```{r}
x <- factor(c('low', 'medium', 'low', 'high', 'low', 'medium'),
            levels = c('low', 'medium', 'high'))
table(x)
unclass(x)
```

## Missing Values

there are two types of missing values\
(-) NA\
(-) NaN\

NaN should to be NA but not vice versa.

```{r}
x <- c(1, 2, NaN, 4)
is.na(x)
is.nan(x)
```

reverse couldn't be applied

```{r}
x <- c(1, 2, NA, 4)
is.na(x)
is.nan(x)
```

## Vextors, Lists, and Data Frames

vector -\> `c()`\
list -\> `list()`\
data frame -\> `data.frame()`\

the differences between thoose function

```{r}
#vector with predifined indices
vct <- c(a=1, b=2, c=3, d=4)
vct

#lists with predefined indices and vector within it partially
lst <- list(a=c(1,4,1), b=2, c=3)
lst

#data frame with list and vector inside
dtfrm <- data.frame(a=list(1,2,3), b=c(2,4,5), c=3)
dtfrm
```

list **could** contain diference number of elements in every index.\
data frame **could not** contain difference number of elements i.e dimension in every column have to be same.\

## dput() and dump()

`dput()` and `dump()` are used to store object

```{r}
y <- data.frame(a=1, b='a')
dput(y)

dput(y, file = 'y.R')
new.y <- dget('y.R') # call dput's file with dget()
new.y
```

store multiple object using `dump()`

```{r}
x <- 'halo'
y <- data.frame(a= c(3,4), b='yes')
z <- c(T, F, T)

dump(c('x', 'y', 'z'), file = 'objdump.R') # putting objects in dump()

rm(x, y, z) # removing objects from the environment

# the object can still be called again using source() although it is no longer in the environment
source('objdump.R') # and now the objects return to the environment
```

## Subsetting Lists

```{r}
x <- list(a = list(c(1, 2, 9), 3, 5), b=c(3, 2, 9), c = 'a')

x[c(1, 1)]
x[[c(1, 1)]]

x[c(1,3)]
x[[c(1,3)]]

x[[1]][[3]]
```

## Partial Matching

```{r}
x <- list(col1 = 1:5, dol1 = c(2, 7, 5), cam = 'd')

x$d # it return an output because index begin with letter 'd' only one
x$c # it doesn't return an output because index begin with letter 'a' more than one

x[['d', exact = FALSE]]
x[['a', exact = FALSE]]
```

## Removing NA values

```{r}
x <- c(NA, 1, 2, 4, NA, 9)
nul <- is.na(x)

x_clean <- x[!nul]
x_clean
```

in multiple objects

```{r}
x <- c(1, 5, NA, 4, NA, 9, 3)
y <- c(1, 5, 4, NA, 9, 3, NA)
clean <- complete.cases(x, y) #complete_cases() used to pair objects with no missing values in each other
x[clean]
y[clean]
```

`complete.cases()` good to use for data frame

# Week 2

## For Looping

```{r}
x <- c('a', 'b', 'c')

for (i in 1:3) {
  print(x[i])
}

for (i in seq_along(x)) {
  print(x[i])
}

for (letter in x) {
  print(letter)
}

for (i in 1:3) print(x[i])
```

nested for loop
```{r}
x <- matrix(1:6, 2, 3, byrow = T)

for(i in seq_len(nrow(x))) {
  for(j in seq_len(ncol(x))) {
    print(x[i, j])
  }
}
```
## While Loop
```{r}
z <- 10

while(z>=3 & z<=10) {
  print(z)
  coin <- rbinom(1, 1, 0.5)
  
  if(coin == 1) {
    z <- z+1
  } else {
    z <- z-1
  }
}
```

## Creat Function
```{r}
x <- matrix(1:20, 4, 5, byrow = T)

countmean <- function(x, removeNA = T) {
  nc <- ncol(x)
  means <- numeric(nc)
  for(i in 1:nc) {
    means[i] <- mean(x[, i], na.rm = removeNA)
  }
  means
}
```