---
title: "JHUDSS C2"
author: "mdzakyjaya"
date: '2022-08-02'
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Week 1 Data Types

## Matriks

pembuatan matriks dapat menggunakan pendefinisan dimensi dengan vector

```{r}
m <- 1:10
dim(m) <- c(2,5)
m
```

pembuatan matriks bisa juga menggunakan funsi cbind() dan rbinf()

```{r}
x <- c(1,3,4)
y <- c(3,4,1)
cmatriks <- cbind(x,y) # menyusun berdasarkan kolom
rmatriks <- rbind(x,y) # menyusun berdararkan baris
```

penamaan dimensi matriks
```{r}
m <- matrix(c(1,2,3,4), nrow = 2, ncol = 2)
dimnames(m) = list(c('a','b'),c('c','d'))
m
```

## Factors

```{r}
x <- factor(c('laki', 'perempuan', 'laki', 'laki', 'perempuan'))
x
table(x) # membuat tabel berisi frequensi dari masing2 faktor
unclass(x) # melihat faktor dari masing value
```

R will order string alpabhetically if ordered levels not predefined.\
defining factors with ordered levels

```{r}
x <- factor(c('low', 'medium', 'low', 'high', 'low', 'medium'),
            levels = c('low', 'medium', 'high'))
table(x)
unclass(x)
```

## Missing Values

there are two types of missing values\
(-) NA\
(-) NaN\

NaN should to be NA but not vice versa.

```{r}
x <- c(1, 2, NaN, 4)
is.na(x)
is.nan(x)
```

reverse couldn't be applied

```{r}
x <- c(1, 2, NA, 4)
is.na(x)
is.nan(x)
```

## Vextors, Lists, and Data Frames

vector -\> `c()`\
list -\> `list()`\
data frame -\> `data.frame()`\

the differences between thoose function

```{r}
#vector with predifined indices
vct <- c(a=1, b=2, c=3, d=4)
vct

#lists with predefined indices and vector within it partially
lst <- list(a=c(1,4,1), b=2, c=3)
lst

#data frame with list and vector inside
dtfrm <- data.frame(a=list(1,2,3), b=c(2,4,5), c=3)
dtfrm
```

list **could** contain diference number of elements in every index.\
data frame **could not** contain difference number of elements i.e dimension in every column have to be same.\