---
title: "JHUDSS C3 Getting and Cleaning Data"
author: "mdzakyjaya"
date: "2022-08-08"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Week 1

## Checking and Creating Directories

```{r}
if (!file.exists('C:/Users/Dtech/Desktop/jhudsps')) {
  dir.create('C:/Users/Dtech/Desktop/jhudsps')
}
```

```{r}
library(jsonlite)
data <- fromJSON('https://api.github.com/users/jtleek/repos')
names(data)
```

# data.table

```{r}
download.file('https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06hid.csv', destfile = 'E:/usc.csv')
data <- read.csv('E:/usc.csv')
```

read data into `data.table`

```{r}
dt <- data.table(data)
dt[VAL==24, .N] #count records with VAL = 24
```

# Week 2

## Reading from MySQL

```{r}
install.packages("RMySQL")
library(RMySQL)

database <- dbConnect(MySQL(), user = 'genome',
                      host = 'genome-mysql.cse.ucsc.edu')

dbGetQuery(database, 'show databases;'); dbDisconnect(database)

dbGetQuery(database, 'use hg19')

length(dbListTables(database)) # count number of tables in corrsepomding database

dbListFields(database, 'HInv') # list attributes in table

hinv <- dbReadTable(database, 'HInv') # copy data frame from database table

dbGetQuery(database, 'select count(*) as records from HInv')

## using sendquery()
query <- dbSendQuery(database, 'select * from HInv')
fetch(query, n=10)
```

# Week 3

## Subsetting & Sorting

```{r}
x <- data.frame('var1'=sample(1:5), 'var2'=sample(6:10), 'var3'=sample(11:15))
x <- x[sample(1:5),] ; x$var2[c(1,3)] = NA
x
```

Logical ands and ors

```{r}
x[(x$var1 <= 3 & x$var3 >=13),] # subsetting with logical and

x[(x$var1 <=3 | x$var3 < 13),] # subsetting with logical or
```

Sorting & Ordering

```{r}
#sorting for column
sort(x$var1)

sort(x$var2, decreasing = T, na.last = T)

#sorting for dataframe
x[order(x$var2, na.last = T, decreasing = T),]
```

Ordering with `plyr`

```{r}
#install.packages('plyr')
library(plyr)

plyr::arrange(x, x$var1)
plyr::arrange(x, desc(var3))
```

Adding Rows and Columns

```{r}
x$var4 <- rnorm(5)
x

# adding by columns
y <- cbind(x, rnorm(4), byrow = T)
y

# adding by rows
z <- rbind(y, rnorm(6))
z
```

## Summarizing Data

importing data

```{r}
restdata <- read.csv('Restaurants.csv')
```

quick overview of the data

```{r}
head(restdata, 3)

tail(restdata, 3)

summary(restdata)

str(restdata)
```

checking quantiles of quantitative variables

```{r}
quantile(restdata$x_coord)

quantile(restdata$x_coord, probs = c(0.2, 0.6, 1))
```

make table for factor variables to quick overview the frequencies of each factors.

```{r}
table(restdata$zipcode, useNA = 'ifany') # useNA will input NA value if any

table(restdata$zipcode, restdata$loc_type) # count frequencies in 2 dimensional table
```

Checking for missing values

```{r}
sum(is.na(restdata$address)) # sum up missing values

any(is.na(restdata)) # checking if there are any missing values in the data

any(is.na(restdata$name)) # checking if there are any missing values in the specific column

all(restdata$loc_type == 'ST') # checking if all data are meet the specific condition

colSums(is.na(restdata)) # to sum up specific boolean criteria in the data

all(colSums(is.na(restdata)) == 0)
```

Checking values with specific characteristics

```{r}
table(restdata$zipcode %in% c(21201, 21287))

# filtering data like where clause in SQL
restdata[restdata$zipcode %in% c(21234, 21239), c('name', 'zipcode', 'city', 'state', 'address')]

# it will give error outcome
restdata[restdata$zipcode %in% c(21234, 21239), c(restdata$name, restdata$zipcode, restdata$city, restdata$state, restdata$address)]

# it's not proper syntax
restdata['zipcode' %in% c(21234, 21239), c('name', 'zipcode', 'city', 'state', 'address')]
```

Cross Tabs

```{r}
data("UCBAdmissions")

# input as dataframe
ucbdf <- as.data.frame(UCBAdmissions)

summary(ucbdf)

# make the cross tab
ct <- xtabs(Freq ~ Admit + Gender, data = ucbdf)

# make the cross tab for all variables
ctall <- xtabs(Freq ~ ., data = ucbdf)
```

Making Flat Tables from Cross Tabs

```{r}
ftable(ctall)
```

Checking size of the data set

```{r}
object.size(ucbdf)

print(object.size(ucbdf), units = "Kb")
```

## Creating New Variables

creating sequences for index for data set

```{r}
s1 <- seq(1, 10, by=2)

s2 <- seq(1, 10, length = 3)

v <- c(1, 6, 7, 10, 3)
s3 <- seq(along = v) # creating indices for vector v
```

subsetting variables

```{r}
restdata$aroundarea <- restdata$zipcode %in% c(21287, 21237, 21234)
table(restdata$aroundarea)
```

creating binary variables

```{r}
restdata$zcnum <- as.integer(restdata$zipcode)
restdata$wrongzip <- ifelse(restdata$zcnum<0, T, F)
table(restdata$wrongzip)
```

creating categorical variables

```{r}
restdata$groupzip <- cut(restdata$zcnum, breaks = quantile(restdata$zcnum, na.rm = T))
table(restdata$groupzip)

table(restdata$zipcode, restdata$groupzip)
```

easier cutting to make categorical variables

```{r}
# install.packages("Hmisc")
restdata$zipgroupease <- Hmisc::cut2(restdata$zcnum, g = 4)
table(restdata$zipgroupease)
```

creating levels of factor variables

```{r}
yesno <- sample(c('yes', 'no'), size = 10, replace = T)
lefact <- factor(yesno, levels = c('yes', 'no'))
lefact
```

create new variables with simultaneously added into new dataframe

```{r}
restdata2 <- plyr::mutate(restdata, zipgroup = Hmisc::cut2(zcnum, g = 4))
table(restdata2$zipgroup, restdata2$zipcode)
```

## Reshaping Data

start with reshaping

```{r}
# install.packages("reshape2")
# library(reshape2)

data("mtcars")

mtcars$carname <- rownames(mtcars)
```

melting dataframe

```{r}
carmelt <- melt(mtcars, id = c('carname', 'gear', 'cyl'), measure.vars = c('mpg', 'hp'))

head(carmelt, 10)
```

casting data frame

```{r}
cyldata <- dcast(carmelt, cyl ~ variable, sum)
```

grouping value with `tapply()`

```{r}
data("InsectSprays")
tapply(InsectSprays$count, InsectSprays$spray, mean)
```

split (grouping without aggregate function)

```{r}
split(InsectSprays$count, InsectSprays$spray)

split(mtcars$carname, mtcars$cyl)
```

using `plyr` package

```{r}
ddply(InsectSprays, 'spray', summarise, jumlah = sum(count))
```

## Using `dplyr` to Easily Dealing with Data Frame

load data
```{r}
library(dplyr)
chicago <- readRDS('chicago.rds')
```

select funtion
```{r}
# select column
head(select(chicago, c(city:date)))

# excluding column
head(select(chicago, -c(city:date)))
```

filter function
```{r}
head(filter(chicago, o3tmean2 > 30 & no2tmean2 < 20))
```

arrange funtion
```{r}
head(arrange(chicago, tmpd))

head(arrange(chicago, desc(date)))
```

rename funtion
```{r}
chicago <- rename(chicago, pm25 = pm25tmean2, dewpoint = dptp, 
                  pm10 = pm10tmean2, o3 = o3tmean2, no2 = no2tmean2)
head(chicago)
```

mutate funtion
to create new variable
```{r}
chicago <- mutate(chicago, diff = no2-o3)
head(chicago)

chicago <- mutate(chicago, tempcat = factor((tmpd > 35),
                                            labels = c('cold', 'hot')))
```

group by funtion
```{r}
hotcold <- group_by(chicago, tempcat)
summarize(hotcold, tmpd = mean(tmpd, na.rm = T))
```
```{r}
chicago <- mutate(chicago, year = as.POSIXlt(date)$year + 1900)
byyear <- group_by(chicago, year)

summarize(byyear, tmpd = mean(tmpd, na.rm = T))
```

using pipeline operator
```{r}
chicago %>% mutate(month = as.POSIXlt(date)$mon + 1) %>% 
  group_by(month) %>% summarize(tmpd = mean(tmpd, na.rm = T))
```

using join in plyr packages
```{r}
df1 <- data.frame(id = c(1:10), x = rnorm(10))
df2 <- data.frame(id = c(1:10), y = rnorm(10))

join(df1, df2)

df3 <- data.frame(id = c(1:10), z = rnorm(10))

join_all(list(df1, df2, df3))
```












